
Reproducible Research: Peer Assessment 1
========================================

## Loading and preprocessing the data

```{r}
#fileURL <- "https://github.com/micbain/RepData_PeerAssessment1/blob/master/activity.zip?raw=true"
zipfile="activity.zip"
#download.file(fileURL, destfile=zipfile)
unzip(zipfile)
activity <- read.csv("activity.csv")
```

## What is the mean total number of steps taken per day?
1. Make a histogram of the total number of steps taken each day

```{r}
str(activity)
summary(activity)



steps.date <- aggregate(steps ~ date, data=activity, FUN=sum)
barplot(steps.date$steps, names.arg=steps.date$date, xlab="Date", ylab="# of steps")
steps.date
```

2. Calculate and report the **mean** and **median** total number of steps taken per day

```{r}
mean(steps.date$steps)
median(steps.date$steps)
```

## What is the average daily activity pattern?
1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
steps.interval <- aggregate(steps ~ interval, data=activity, FUN=mean)
plot(steps.interval, type="l",  ylab="averaged across all days (y-axis)", xlab="average number of steps taken")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
steps.interval$interval[which.max(steps.interval$steps)]
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)

```{r}
sum(is.na(activity))
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc


5-minute intervals is used as fillers for missing values.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activity <- merge(activity, steps.interval, by="interval", suffixes=c("",".y"))
missingsteps <- is.na(activity$steps)
missing <- subset(activity, is.na(activity$steps))
summary(missing$date)
activity$steps[missingsteps] <- activity$steps.y[missingsteps]
activity <- activity[,c(1:3)]
```

4. Make a histogram of the total number of steps taken each day and calculate and report the **mean** and **median** total number of steps taken per day. 

Do these values differ from the estimates from the first part of the assignment? 

What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
steps.date <- aggregate(steps ~ date, data=activity, FUN=sum)
barplot(steps.date$steps, names.arg=steps.date$date,  xlab="Date", ylab="# of steps")
mean(steps.date$steps)
median(steps.date$steps)
```

###Answer: The impact of the missing data seems low, at least when estimating the total number of steps per day. Apparently all the data for 8 of the 61 days of observation is missing, which explains the consistency of the number of missing data per time interval.


## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is aweekday or weekend day.

```{r, cache=TRUE}
daytype <- function(date) {
     if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
          "weekend"
     } else {
          "weekday"
     }
}
activity$daytype <- as.factor(sapply(activity$date, daytype))
```

2. Make a panel plot containing a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
par(mfrow=c(2,1))
for (type in c("weekend", "weekday")) {
steps.type <- aggregate(steps ~ interval, data=activity, subset=activity$daytype==type, FUN=mean)
plot(steps.type, type="l", main=type,  xlab="Steps (5mins)", ylab="Average steps taken (all days)")
}
```
